---
title: "In-Class Lab 2"
author: "Javier Patrón"
date: "r Sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(LexisNexisTools)
library(dplyr)
library(readr)
library(stringr)
library(here)
library(tidytext)
library(tidyr)
library(ggplot2)
```

Sentiment analysis is a tool for assessing the mood of a piece of text. For example, we can use sentiment analysis to understand public perceptions of topics in environmental policy like energy, climate, and conservation.

### Introduction to the example

Today's example will use data from the Nexis Uni (formerly Lexis Nexis) database, accessed through the UCSB library. There are a large number of news and other full-text publications available through the database. I'm interested in the discussion around deforestation of the Amazon rainforest. The current president of Brazil, Luiz Inacio Lula da Silva recently completed his first 100 days in office after running on a pledge to reduce the rate of deforestation. I'd like to know how the sentiment in coverage of deforestation changed after his election.

```{r raw_data}


my_files <- list.files(pattern = ".docx", path = getwd(),
                       full.names = TRUE, recursive = TRUE, ignore.case = TRUE)

class_files <- readRDS(here("my_files.RDS"))
dat <- readRDS(here("lexisdat.RDS"))

```

We'll use the {LexisNexisTools} package to handle the documents from our Nexis search.

```{r}
#dat <- lnt_read(my_files)

# The @ is for indexing within tibbles
meta_df <- dat@meta
articles_df <- dat@articles
paragraphs_df <- dat@paragraphs

dat2 <- tibble(Date = meta_df$Date, 
              Headline = meta_df$Headline, 
              id = dat@articles$ID, 
              text = dat@articles$Article)

```

```{r get_bing}

# create a list of all the words in the diccionary and if they have a sentiment word attached like "positive" or "negative".

bing_sent <- get_sentiments('bing') #grab the bing sentiment lexicon from tidytext
head(bing_sent, n = 20)
```

1.  Score words using bing lexicon
```{r text_words}
# Create a new data frame that has a row for each word
text_words <- dat2 |> 
  unnest_tokens(output = word, 
                input = text,
                token = 'words')

# Create a new data frame that has 
sentiment_words <- text_words |> 
  anti_join(stop_words) |> # Removing the unuseful words
  inner_join(bing_sent)  |> # Adding the sentiment related column
  mutate(sent_num = case_when(sentiment == "negative" ~ -1,
                              sentiment == "positive" ~ 1))

```

2.  Calculate mean sentiment (by word polarity) across articles
```{r mean_sent}
sentiment_article <- sentiment_words |> 
  group_by(Headline) |> 
  count(id, sentiment) |> 
  pivot_wider(names_from = sentiment, 
              values_from = n,
              values_fill = 0) |> 
  mutate(polarity = positive - negative)

mean(sentiment_article$polarity)

```

3.  Sentiment by article plot

Let's try a very basic plot to see the amount of positive vs. negative articles.
```{r plot_sent_scores}

ggplot(sentiment_article, aes(x = id)) +
  theme_classic() +
  geom_col(aes(y = positive), stat = "identity", fill = "slateblue4") +
  geom_col(aes(y = negative), stat = "identity", fill = "red") +
  theme(axis.title.y = element_blank()) +
  labs(title = "Sentiment Analysis",
       subtitle = "Amazon Deforestation", 
       y = "Sentiment score")

```

##### Origin of the NRC lexicon

"These guys selected about 10,000 words from an existing thesaurus... and then created a set of five questions to ask about each word that would reveal the emotions and polarity associated with it. That's a total of over 50,000 questions.

They then asked these questions to over 2,000 people, or Turkers, on Amazon's Mechanical Turk website, paying 4 cents for each set of properly answered questions.

The result is a comprehensive word-emotion lexicon for over 10,000 words."

4.  nrc emotion words

Let's take a look at the most common emotion words in the data set

```{r nrc_sentiment}
# Create a new data frame with the sentiments per word.
nrc_sent <- get_sentiments("nrc")

# Create a new data frame that has the text_words with the column of nrc sentiment
nrc_word_count <- text_words |> 
  anti_join(stop_words, by= "word") |> 
  inner_join(nrc_sent) |> 
  count(word, sentiment, sort = T) 
```


Let's break it out and plot the contributions by particular emotion categories.

```{r sent_counts}
sent_counts <- text_words |> 
  anti_join(stop_words, by = "word") |> 
  group_by(id) |> 
  inner_join(nrc_sent) |> 
  group_by(sentiment) |> 
  count(word, sentiment, sort = T)

sent_counts |> 
  group_by(sentiment) |> 
  slice_max(n, 
            n= 5) |> 
  ungroup() |> 
  mutate(word = reorder(word,n)) |> 
  ggplot(aes(n, word, 
             fill = sentiment)) +
  geom_col(show.legend = F) +
  facet_wrap(~sentiment, 
             scales = "free_y") +
  labs(x= "Contribution to Sentiment", 
       y = NULL)
  
  
```

Now let's do a quick comparison to articles from the 100-days leading up to the beginning of Lula's term.

```{r}
setwd(here("Nexis/Pre"))

pre_files <- list.files(pattern = ".docx", path = getwd(),
                       full.names = TRUE, recursive = TRUE, ignore.case = TRUE)

pre_dat <- lnt_read(pre_files)

pre_meta_df <- pre_dat@meta
pre_articles_df <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

pre_dat2<- tibble(Date = pre_meta_df$Date, Headline = pre_meta_df$Headline, id = pre_dat@articles$ID, text = pre_dat@articles$Article)

```

```{r pre_text_words}
text_words <- pre_dat2  %>%
  unnest_tokens(output = word, input = text, token = 'words')
 
 sent_words <- text_words%>% #break text into individual words
  anti_join(stop_words, by = 'word') %>% #returns only the rows without stop words
  inner_join(bing_sent, by = 'word') #joins and retains only sentiment words
```

```{r mean_pre}
pre_sentiment <- sent_words %>%
  count(id, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(polarity = positive - negative)
```



